---
title: "Bisphosphonates adverse-events network meta-analysis: event-specific results"
author: ""
date: "`r Sys.Date()`"
output:
 html_document:
  code_folding: hide
---
[Back to Appendix 2](../app2_index.html)
```{r, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      fig.width=15,
                      fig.height=15)
library(knitr)
library(dplyr)
filter <- dplyr::filter
library(flextable)
library(rlang)
library(ggplot2)
library(gridExtra)
library(plotly)
setwd("~/work/winton/adverse/vignettes")
load_all("..")
events <- names(models_all)
nev <- length(events)
load(file="ors.rda")
```	
## Irregular menstruation

```{r, warning=FALSE}
arrfn("IRREGULAR MENSTRUATION")
```

### Network geometry summary

Proportion of potential comparisons with direct data: 0.5

Proportion of treatments compared to more than one other: 0.8

Median number of studies per direct comparison: 1 (interquartile range 1-1)

Proportion of direct comparisons informed by more than one study: 0

## Meta analysis results
Estimates (and 95% credible intervals) of the odds ratio of each bisphosphonate treatment, versus observation only, under

* (green) the best-fitting network meta-analysis model (if feasible)

* (blue) an equivalent meta-analysis of direct comparisons of the same treatment versus an observation-only control, and

* (red) all study-specific direct comparisons odds ratios corresponding to direct comparisons of the treatment with a no-treatment control.  Placebo controls (dotted lines) are distinguished from observation-only controls (solid lines).

The panels distinguish the treatment categories under the best-fitting network meta-analysis model.

Note:

* there is one red line for each active study arm in each study containing a placebo or observation-only control.  Note that this excludes comparisons of one bisphosphonate against another, which contribute to the network-meta analysis pooled estimates along with the direct comparisons.  

* the blue line (direct-data pooled estimate) is the weighted average of all solid red lines (study-specific comparisons against observation-only).  If there is no blue line on a particular panel, there were no studies that directly compared that treatment against an observation-only control.

* the green line (network meta-analysis pooled estimate) will not necessarily appear to be the average of the red lines, since the network meta-analysis also includes indirect data from the bisphosphonate-bisphosponate comparisons that are not illustrated 

* there is no green line if no network meta-analysis was feasible.

```{r, warning=FALSE}
plotarr[[56]]$ests

```


## Comparison of different treatment groupings


Network meta-analysis models defined on different groupings of treatments are compared.

*  The treatment groupings compared for this symptom are those for which there is a connected network of trials reporting that symptom.

* Lines indicated pooled estimates of the odds ratio for each treatment against observation only, under the network meta-analysis model on that treatment grouping.

* The statistical fit of the models on different treatment groupings is compared using the deviance information criterion (DIC).  DIC are presented relative to the best-fitting model (DIC = 0), and higher DIC indicate worse fit.

For the event IRREGULAR MENSTRUATION, observation-only and placebo are considered separately.

```{r, warning=FALSE}
plotarr[[56]]$modcomp

```


